---
title: "Session 1: Daily Usage Data"
output: prettydoc::html_pretty
date: '2023-07-13'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(readxl)

```


## The data

The data we're using today are new experimental statistics looking at the percentage of taxis and PHVs (private hire vehicles) of different ages by region.

The data is available in the data folder associated with this project ("Data/taxi0116.xlsx") in a non-tidy, human-readable format, and you can see the first rows of the data below:

```{r}
readxl::read_excel("../Data/taxi0116.xlsx", sheet = 3, skip = 5) %>%
  head() %>%
  knitr::kable()
```

The data covers both taxis and PHVs in separate tables. Each different age range of the vehicles is recorded in a separate column, and each row is for a different region. 

Each cell contains a value of percentage of vehicles in each region or country that are in each age bracket. Each region or country sums to 100%.



## The task

1. Read the data in to R. You don't need to save the file locally, you can do this directly from the web link

2. Check that the data is clean, and the different modal names are in a publication-ready format

3. Create a ggplot chart of the data, with dates on the x axis and value on the y axis, with one line per transport mode.

4. Make your chart publication-worthy! Aspects you may want to consider include:

* The theme and colours used in your charts
* The formatting and labelling of your chart axes
* What should the date range of your data be?
* Do you want to include weekends and bank holidays?
* Do you want to include every mode? Could you split the chart to show different groupings on different charts?

