---
title: "Session 2: Google Mobility Data"
output: prettydoc::html_pretty
date: '2023-07-11'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(data.table)
```

```{css, echo=FALSE}
.spoiler {
  visibility: hidden;
}

.spoiler::before {
  visibility: visible;
  content: "HINT! Hover me to see a hint to help on this step."
}

.spoiler:hover {
  visibility: visible;
}

.spoiler:hover::before {
  display: none;
}
```


## The data

The data we're using today is the now-defunct Google Mobility Data Series, produced to allow people to monitor how people travelled in different countries across the world, based on Google Maps data. This data provides a daily figure for change in mobility compared to the baseline, for different countries, regions, and sub-regions around the world. 

The data is available [here](https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv) in a semi-tidy, machine-readable format, and you can see the first 10 rows of the data below:

```{r}
data.table::fread("https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv") %>%
  head() %>%
  knitr::kable()
```

The data covers travel in the following types of area:

* retail_and_recreation 
* grocery_and_pharmacy
* parks 
* transit_stations
* workplaces 
* residential

Part of the challenge of this data is how large it is; you will need to get used to handling it solely R rather than looking at the CSV in Excel, and working efficiently so your code doesn't take ages to run!

## The task

1. Read the data in to R. You don't need to save the file locally, you can do this directly from the web link (https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv) 

2. Filter the data down to just the United Kingdom and the country-level data only. **[The country_region_code for the United Kingdom is GB, and country-level data is indicated by no sub_region_1]{.spoiler}**

3. Bring the data into a tidy format, so there is only one column for value and one for type of area. **[You will want to use the tidyr pivot_longer() function for this]{.spoiler}**

4. Create a ggplot chart of the data, with dates on the x axis and value on the y axis, with one line per type of area.

4. Make your chart publication-worthy! Aspects you may want to consider include:

* The theme and colours used in your charts
* The formatting and labelling of your chart axes
* What should the date range of your data be?
* Do you want to include weekends and bank holidays?
* Do you want to include every type of area? Or would you like to show more than one country, or a breakdown by region? Can you even show multiple types of area _and_ different countries?

