---
title: "Session 1: Daily Usage Data"
output: prettydoc::html_pretty
date: '2023-07-11'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(rvest)
library(dplyr)

url <- "https://www.gov.uk/government/statistics/transport-use-during-the-coronavirus-covid-19-pandemic"

#Set up a function to neatly read the file in
read_online <- function(url){
  #Read all HTML
  pg <- read_html(url)
  #Pull href html links
  pg <- html_attr(html_nodes(pg, "a"), "href")
  
  #Keep only ones which link to Covid CSV and ignore duplicates
  pg <- pg[grepl("[.]csv$", pg)] %>% 
    unique()
  
  if(length(pg) != 1){
    stop("A single CSV file could not be identified at this URL")
  }

  return(pg)
  
}

```


## The data

The data we're using today is the daily transport usage stats, initially produced for Covid reporting. This data provides a daily figure for usage by mode, indexed to a pre-Covid baseline figure.

The data is available [here](`r read_online(url)`) in a tidy, machine-readable format, and you can see the first 10 rows of the data below:

```{r}
data.table::fread(read_online(url)) %>%
  head() %>%
  knitr::kable()
```

The data covers the following modes:

* all_motor_vehicles
* tfl_tube 
* tfl_bus
* national_rail 
* cycling
* bus_excluding_london

Motor vehicles is further broken down by vehicle type:

* cars
* light_commercial_vehicles
* heavy_goods_vehicles

Where no data is available for a specific date, you can treat that as an NA value.


## The task

1. Read the data in to R. You don't need to save the file locally, you can do this directly from the web link (`r read_online(url)`) 

2. Check that the data is clean, and the different modal names are in a publication-ready format

3. Create a ggplot chart of the data, with dates on the x axis and value on the y axis, with one line per transport mode.

4. Make your chart publication-worthy! Aspects you may want to consider include:

* The theme and colours used in your charts
* The formatting and labelling of your chart axes
* What should the date range of your data be?
* Do you want to include weekends and bank holidays?
* Do you want to include every mode? Could you split the chart to show different groupings on different charts?

